A Study on Common Bug Fix Patterns in Rust

### 2022-11-09 methodology.tex
changes added before results

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change---terminals and nonterminals. Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the items of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

@@

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?'' combinatorial explosion}

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example? ``Fig 2?'' introduction to the next section}. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters. We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.

@@

% TODO fix this caption
\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{clusters.pdf}
\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows. In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$. The two orange bars in the right show similar data but for the borrow-checker $D_{b}$ database. In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
\end{figure}

@@

In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes. Also, we prefer reporting clusters that manifest more-specific patterns. An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs. For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Adding a new statement to a function's body':

@@

// Second Pattern
fn foo2() {
    stmt1;
++  stmt2;
}

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster. Next, we read the code and the bug reports related to all the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes. Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.

>>>>

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change---terminals and nonterminals. Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the \textit{items} of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

@@

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields a combinatorial explosion; this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs. In Section~\ref{sec:weighting_scheme}, we provide and example on how this approach works.

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters. We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.
\todo[inline]{Wouldn't the reviewers complain why you haven't put k-means data in the paper?}

@@

% TODO fix this caption
\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{clusters.pdf}
\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows. In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$. The two orange bars in the right show similar data but for the borrow-checker database $D_{b}$. In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
\end{figure}

@@

In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes. Also, we prefer reporting clusters that manifest more-specific patterns. An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs. For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Changing a statement in a function's body':

@@

// Second Pattern
fn foo2() {
--  stmt1;
++  stmt2;
}

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster. Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns. Out of these 428 clusters, 35 clusters included more than 50 points. That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them. Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project pattern. Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters. For each cluster, we read the code and the bug reports of the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 20 of the remaining clusters---the ones that manifested more-specific changes. Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.
